### 2021-08-20 analysis/index.Rmd
further development of manuscript

---
title: "Untitled"
author:
  - "Bjørn Peare Bartholdy"
  - Amanda G. Henry
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    df_print: kable
    toc: no
    latex_engine: xelatex
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: default
  bookdown::html_document2:
    df_print: kable
    keep_md: true
linestretch: 2.0 # double-spacing for review purposes
bibliography: paper/refs.bib
csl: paper/apa.csl
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.path = "./figures/")
library(tidyverse)
library(broom)
library(here)
library(patchwork)
# load R scripts
source(here("analysis/scripts/analysis.R"))
source(here("analysis/scripts/tables.R"))
source(here("analysis/scripts/figures.R"))
## Compile refs.bib using rbbt package ##
file_paths <- here("analysis/paper", c("intro.Rmd", "matmet.Rmd", "disc.Rmd"))
cite_keys <- sapply(file_paths, rbbt::bbt_detect_citations) 
cite_keys <- unlist(cite_keys, use.names = FALSE)
rbbt::bbt_write_bib("./paper/refs.bib", keys = cite_keys, overwrite = T)
## end refs.bib ##
```

```{r intro, child="paper/intro.Rmd"}

```

```{r matmet, child="paper/matmet.Rmd"}

```

```{r res, child="paper/results.Rmd"}

```

```{r disc, child="paper/disc.Rmd"}

```

# Acknowledgements {-}

<!-- If you're reading this, thank you for taking the time to read the source 
documents. You're awesome! -->

# References cited