### 2021-09-30 analysis/paper/abstract.Rmd
revisions to first draft

# Abstract

```{r include=FALSE}
w_count <- wordcountaddin::word_count()
```

<!-- One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline -->

<!-- Two to three sentences of more detailed background, comprehensible to scientists in related disciplines -->

Dental calculus, mineralised plaque, has proven to contain a wealth of information 
on the dietary habits of past populations. The extraction and identification of
starch granules contained within dental calculus has already provided valuable
dietary insights from a wide range of regions and time periods, and seem to preserve 
well in the archaeological record.
<!-- One sentence clearly stating the general problem being addressed by this particular study -->
While the majority of studies have focused on microfossil extraction and
identification from archaeological remains, <!--
as well as methods of extraction and decontamination,--> few studies have 
addressed the biases of starch retention in dental calculus, and have been
limited to *in vivo* studies on modern humans and non-human primates.
<!--
and have primarily been 
conducted on modern humans, non-human primates, and archaeological remains; 
all of which have 
methodological challenges of dealing with living (or once living) organisms, 
and the variables that are difficult to control. <!-- rephrase -->

<!-- One sentence summarising the main result (with the words "here we show" or their equivalent) -->
Here, we present a multispecies oral biofilm model, which allows experimental
research to be conducted on *in vitro* dental calculus in a controlled laboratory
setting. The biofilms were exposed treatment solutions with known quantities of
dietary starches (wheat and potato) during the 25-day growth period. After this,
the starch granules were extracted from the mature biofilm (dissolution in EDTA),
and counted. We 
show that the granule counts extracted from the model dental calculus represented 
a low proportion of the total counts (ranging from 0.064% to 0.161%) exposed to 
the biofilms throughout the experiment. Additionally, we found that the ratios
of granule sizes from the extracted starch granules differed from the original
treatment solutions, with large granules (>20 $\mu$m) consistently being
under-represented.
Additionally, the size ratios of granules within starch species shifted between the input starch solutions and the granules extracted from the model calculus, with large granules (>20 $\mu$m) consistently being underrepresented. We also found a correlation
between the absolute granule counts and dry-weight of the biofilm (r = ), as well
as between the concentration (count per mg) of granules and dry-weight (r = ).

<!-- Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge -->

<!-- One or two sentences to put the results into a more general context -->

<!-- Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. -->
Our results reinforce previous *in vivo* studies suggesting that dental calculus 
presents a very small, and partly biased picture of the original dietary intake 
of starches, with an over-representation of plants producing granules smaller than
20 $\mu$ in size. The experimental model presented here is well-suited to address
the need for further validation of methods and biases associated with dietary
research on dental calculus.

Word count: `r w_count`