### 2022-02-27 analysis/supplementary-materials/supp-mat.Rmd
update supp mat

---
title: "Supplementary Material"
output:
  bookdown::html_document2:
    df_print: paged
    toc: yes
    toc_float: yes
  bookdown::pdf_document2:
    toc: yes
    df_print: kable
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(patchwork)
source(here("analysis/scripts/analysis.R"))
source(here("analysis/scripts/figures.R"))
source(here("analysis/scripts/tables.R"))
source(here("analysis/scripts/amylase-assay.R"))
#raw_data_samples <- read_csv(here("analysis/data/raw_data/starch_counts.csv"))
#raw_data_solutions <- read_csv(here("analysis/data/raw_data/solution_counts.csv"))
```

# R Session info

```{r}
print(sessionInfo(), locale = FALSE)
```

# Experimental setup

```{r plate-lid-1, fig.cap="The 24 deepwell plate and the lid with pegs (substrata)"}
knitr::include_graphics(here("analysis/supplementary-materials/figures/plate_lid_side.jpg"))
```

```{r plate-lid-2, fig.cap="The 24 deepwell plate with the lid (almost) on."}
knitr::include_graphics(here("analysis/supplementary-materials/figures/plate_lid_on.jpg"))
```

# Protocols

All protocols are available on [protocols.io](https://www.protocols.io).

Creating the artificial saliva: <https://www.protocols.io/view/artificial-saliva-bva9n2h6>

Creating the CPMU solution: <https://www.protocols.io/view/cpmu-bv8pn9vn>

Biofilm growth protocol: <https://www.protocols.io/view/biofilm-growth-with-starch-treatment-bu7jnzkn>

Amylase activity assay: <https://www.protocols.io/view/amylase-activity-bw8jphun>

# Raw data

The raw data can be downloaded from OSF:

Solution counts: <https://osf.io/kz3b2/>  
Sample counts: <https://osf.io/kz3b2/>

```sh
# solution counts
wget https://osf.io/kz3b2/download -O solution_counts.csv

# sample counts
wget https://osf.io/kz3b2/download -O starch_counts.csv
```

## Metadata for raw data files

Counts represent the absolute number of starches counted on a slide

**starch_counts.csv**

| variable | description |
|----|----|
| sample | Sample number. |
| plate | Plate number that the sample came from. |
| row | Which row on the plate the sample came from. |
| s | Small starch count. |
| m | Medium starch count. |
| l | Large starch count. |
| total | Sum of s, m, and l. |
| treatment | Treatment solution to which the samples were exposed. |
| starch | Type of starch that was counted. |
| weight | Weight of the biofilm sample. |
| vol | Total volume of EDTA in which the sample was dissolved. |
| portion_slide | Proportion of the microscope slide that was counted. Total transects on slide divided by counted transects. |

**solution_counts.csv**

| variable | description |
|----|----|
| solution | Type of starch in solution. |
| concentration | Concentration (%w/v) of starch in solution. | 
| vol_slide | Volume of solution added to slide. |
| vol_total | Total volume of solution in aliquot. |
| portion_slide | Proportion of slide that was counted. Total transects on slide divided by counted transects. |
| slide | Slide number. |
| starch | Starch type counted. |
| s | Small starch count. |
| m | Medium starch count. |
| l | Large starch count. |
| total | Sum of s, m, and l. |

# Microscope images

```{r img-1, echo=FALSE, fig.cap="Image of starch granules extracted from a potato treatment sample"}
knitr::include_graphics(here("analysis/supplementary-materials/figures/starches_w_bar.jpg"))
```

```{r img-2}
knitr::include_graphics(here("analysis/supplementary-materials/figures/SNAP-103412-0006.jpg"))
```

```{r img-3, fig.cap="Microscope image of wheat starch from a wheat treatment sample."}
knitr::include_graphics(here("analysis/supplementary-materials/figures/SNAP-164650-0012.jpg"))
```

# Amylase activity

Amylase activity in U/mL enzyme, where U is the amount of enzyme needed to
release 1 $\mu$mole maltose from starch in six minutes at 36 &deg;C.

Tables containing the amylase activity results for both plates and both
photometric readings conducted on each plate. Samples (rows) were analysed in
triplicates (columns).

```{r amylase-activity-results, message=FALSE}
# table of results reported in units amylase per mL enzyme (but let's be honest,
  # ...it doesn't really matter what the unit is. No activity is no activity)
cols <- c("1", "2", "3")  # sample triplicates
rows <- c("S1", "S2", "S3", "B1", "B2", "B3", "B4", "B5", "BT1", "BT2", "BT3")
plt1_ph1_result <- rbind(sal1_ph1, bmm1_ph1)
rownames(plt1_ph1_result) <- rows
plt1_ph2_result <- rbind(sal1_ph2, bmm1_ph2)
rownames(plt1_ph2_result) <- rows
plt2_ph1_result <- rbind(sal2_ph1, bmm2_ph1)
rownames(plt2_ph1_result) <- rows
plt2_ph2_result <- rbind(sal2_ph2, bmm2_ph2)
rownames(plt2_ph2_result) <- rows
```

```{r amylase-plt1-read1}
plt1_ph1_result %>%
  knitr::kable("Amylase activity in U/mL for plate 1, photometric read 1.")
```

```{r amylase-plt1-read2}
plt1_ph2_result %>%
  knitr::kable("Amylase activity in U/mL for plate 1, photometric read 2.")
```

```{r amylase-plt2-read1}
plt2_ph1_result %>%
  knitr::kable("Amylase activity in U/mL for plate 2, photometric read 1.")
```

```{r amylase-plt2-read2}
plt2_ph2_result %>%
  knitr::kable("Amylase activity in U/mL for plate 2, photometric read 2.")
```

# Control samples

```{r}
raw_counts %>%
  filter(treatment == "control") %>%
  select(!c(vol, portion_slide, s, m, l))
```

Only the total starch count was considered for control samples, as size was
deemed irrelevant.

# Count corrections

Slide transects were calculated by counting the number of transects on the 
cover slip under the microscope. This was done by starting in the bottom-left
corner, and counting the total number of full fields-of-view across the cover 
slip to the bottom-right corner. The total number of transects was 29 (verified
multiple times).

A 1 mL aliquot of each of the original treatment solutions was taken, from which
10 $\mu$L was taken and mounted on a microscope slide and mixed with 10 $\mu$L 
20% (v/v) glycerol. Solution counts were extrapolated from a slide (10 $\mu$L) 
to the quantity in a 1 ml solution,
and then multipled by 16 days to achieve the total number of granules that were
exposed to the samples:

$$ \text{corrected count} = \text{raw count} \times \frac{\text{total slides}}{\text{counted slides}} \times 100 \mu L \times 16 \text{ days} $$

Samples were submerged in 50--100 $\mu$L EDTA, from which 20 $\mu$L was mounted 
on a microscope slide ($V_{slide}$) and counted. Sample counts were extrapolated to the full
volume of EDTA ($V_{sample}$) in which the sample was submerged (i.e. 50--100 $\mu$L).

$$ \text{Corrected count} = \text{raw count} \times (\text{portion of slide})^{-1} \times \frac{V_{sample}}{V_{slide}} $$

# Some additional plots

Bar plot for the total count of granules exposed to the samples over the
duration of the experiment,

```{r solution-bar-pl}
sol_long %>%
  filter(size == "total") %>%
  group_by(treatment, starch) %>%
  ggplot(aes(x = treatment, y = count, fill = treatment, col = starch)) +
  geom_col(size = 1.5) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey"),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank()) +
  scale_fill_viridis_d() +
  scale_color_viridis_d(begin = 0.5)
```

and box plot with superimposed points (with added jitter) for the extrapolated
mean counts of granules extracted from the samples.

```{r sample-box-plot}
corr_comb %>%
  filter(treatment != "control") %>%
  ggplot(aes(x = treatment, y = total, 
             shape = treatment)) +
    geom_boxplot(aes(fill = treatment), alpha = 0.5) +
    geom_jitter(aes(col = treatment), width = 0.3, size = 2) +
    scale_color_viridis_d() +
    theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey"),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank()) + # remove y-axis title
    scale_fill_viridis_d()
```

Extracted-granule counts separated by treatment and size, including error bars:

```{r counts-bar, fig.cap="l = large, m = medium, s = small."}
corr_counts_long %>%
  filter(size != "total",
         treatment != "control") %>%
  group_by(treatment, starch, size) %>%
  summarise(sd = sd(count, na.rm = T),
            count = mean(count, na.rm = T)) %>%
  #mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = count, fill = size)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = count, ymax = count + sd), width = 0.2, position = position_dodge(0.9)) +
    facet_wrap(~ treatment, scales = "free") +
    scale_fill_viridis_d() +
    theme_bw()
```

Size distribution (in %) within the solutions (top) and samples (bottom):

```{r size-ratios-bar, fig.cap="l = large, m = medium, s = small.", message=FALSE, warning=FALSE}
sol_size_pl <- sol_corr %>%
  group_by(solution, starch) %>%
  summarise(across(c(s, m, l, total), mean, na.rm = T)) %>%
  pivot_longer(cols = c(s,m,l, total), values_to = "count", names_to = "size") %>%
  filter(size != "total") %>%
  group_by(solution, starch) %>%
  mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = percent, fill = size)) +
    geom_col(position = "dodge") +
    facet_wrap(~ solution, scales = "free_x") +
    scale_fill_viridis_d() +
    theme_bw() +
    labs(x = "")

samp_size_pl <- corr_counts_long %>%
  filter(size != "total",
         treatment != "control") %>%
  group_by(treatment, starch, size) %>%
  summarise(count = mean(count, na.rm = T)) %>%
  mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = percent, fill = size)) +
    geom_col(position = "dodge") +
    facet_wrap(~ treatment, scales = "free_x") +
    scale_fill_viridis_d() +
    theme_bw()

sol_size_pl / samp_size_pl + plot_layout(guides = "collect")
```

Separated correlation plots. These are the same plots as in the main paper, just larger.

```{r cor-plot, warning=FALSE, fig.cap="Scatter plot of sample weight and standardised starch count by z-score for separated treatments."}
pl_cor
```

```{r cor-plot2, warning=FALSE, fig.cap="Scatter plot of sample weight in mg and standardised count of starch grains per mg calculus."}
pl_cor2
```

## ...and a table

Differences in size ratios (%) of granules between the solutions and the samples. 
Negative values indicate a loss of granules from solution to sample.

```{r diff-tbl}
size_diff %>%
  mutate(across(where(is.numeric), signif, 3))
```