### 2022-02-28 analysis/paper/abstract.Rmd
add S prefix to all tables and figs in supp mat

---
output:
  bookdown::pdf_document2:
    keep_tex: true
---

# Abstract

```{r include=FALSE}
library(here)
library(patchwork)
source(here("analysis/scripts/analysis.R"))
source(here("analysis/scripts/tables.R"))
perc_tbl <- prop %>%
  ungroup() %>%
  select(!c(sample, solution, prop)) %>%
  pivot_wider(names_from = size, values_from = perc) %>%
  add_row(perc_comb_row, .after = 2)
# prepare cor output for reporting
cor_r <- signif(count_cor$estimate, 3)
cor_ci <- signif(count_cor$conf.int, 3)
cor2_r <- signif(starch_cor$estimate, 3)
cor2_ci <- signif(starch_cor$conf.int, 3)

#w_count <- wordcountaddin::word_count()
```

<!-- One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline -->

<!-- Two to three sentences of more detailed background, comprehensible to scientists in related disciplines -->

<!-- One sentence clearly stating the general problem being addressed by this particular study -->

<!-- One sentence summarising the main result (with the words "here we show" or their equivalent) -->

<!-- Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge -->

<!-- One or two sentences to put the results into a more general context -->

<!-- Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. -->

Dental calculus has proven to contain a wealth of information 
on the dietary habits of past populations. These insights have, to a large extent,
been obtained by the extraction and identification of
starch granules contained within the mineralised dental plaque from a wide range 
of regions and time periods. The scope of previous studies have been limited to 
microfossil extraction and identification to reconstruct dietary preferences from
the archaeological record, and few studies have attempted to address the biases 
of starch retention in dental calculus. Those that have considered this problem 
have been limited to *in vivo* studies on modern humans and non-human primates.
Here, we present a multispecies oral biofilm model, which allows experimental 
research on  starch incorporation and retention to be conducted on *in vitro* 
dental calculus in a controlled laboratory setting. The biofilms were exposed 
to treatment solutions with known quantities of
dietary starches (wheat and potato) during the 25-day growth period. After this,
the starch granules were extracted from the mature biofilm (by dissolution in EDTA),
and counted. We show that the granule counts extracted from the model dental calculus 
represented a low proportion (ranging from `r paste(min(perc_tbl$total), max(perc_tbl$total), sep = " to ")`) of the total number 
of granules exposed to the biofilms throughout the experiment.
Additionally, we found that the ratios of granule sizes from the extracted 
starch granules differed from the original treatment solutions, with large 
granules (>20 $\mu$m) consistently being under-represented.
We also found a correlation
between the absolute granule counts and dry-weight of the biofilm 
(*r* = `r cor_r`, 90%CI[`r paste(cor_ci[1], cor_ci[2], sep = ", ")`]), as well
as between the concentration (count per mg) of granules and dry-weight 
(*r* = `r cor2_r`, 90%CI[`r paste(cor2_ci[1], cor2_ci[2], sep = ", ")`]).  
Our results reinforce previous *in vivo* studies suggesting that dental calculus 
presents a very small, and partly biased picture of the original dietary intake 
of starches, with an over-representation of plants producing granules smaller than
20 $\mu$m in size. The experimental model presented here is well-suited to address
the need for further validation of methods and biases associated with dietary
research on dental calculus.